\newpage
\section{Master Configuration Guide}

To deploy a robust, universal AI agent toolkit, you can combine the best servers from both the Node (npx) and Python (uvx) ecosystems into a single master configuration block. 

When functionalities overlap (such as the \lstinline[style=inline]|fetch| capability), 
the Node implementation (\lstinline[style=inline]|@modelcontextprotocol/server-fetch|) is 
typically chosen due to its widespread adoption and canonical support, 
while \lstinline[style=inline]|uvx| is preferred for native local system tasks 
(like file I/O and Git history).

\subsection{The Universal JSON Profile}
This configuration aggregates repository control, system auditing, web discovery, 
and reasoning. Many modern AI clients (like the Gemini CLI) support expanding local 
environment variables directly within the JSON block using the 
\lstinline[style=inline]|${ENV_VAR_NAME}| syntax, ensuring you don't leak secrets into your config files. 

\textit{Note: If a referenced environment variable is not defined on your local machine, 
the AI client will typically fail to start that specific MCP server (often logging a connection or authentication error), 
while the rest of the correctly configured servers will continue to function normally. 
Ensure these variables are exported in your shell profile 
(e.g., \lstinline[style=inline]|.bashrc| or \lstinline[style=inline]|.zshrc|) before launching the client.}

\begin{lstlisting}[style=json]
{
  "mcpServers": {
    "github": {
      "command": "npx",
      "args": ["-y", "@modelcontextprotocol/server-github"],
      "env": { "GITHUB_PERSONAL_ACCESS_TOKEN": "${GITHUB_PERSONAL_ACCESS_TOKEN}" }
    },
    "git-history": {
      "command": "uvx",
      "args": ["mcp-server-git"]
    },
    "local-filesystem": {
      "command": "uvx",
      "args": ["mcp-server-filesystem", "/"]
    },
    "fetch": {
      "command": "npx",
      "args": ["-y", "@modelcontextprotocol/server-fetch"]
    },
    "brave-search": {
      "command": "npx",
      "args": ["-y", "@modelcontextprotocol/server-brave-search"],
      "env": { "BRAVE_API_KEY": "${BRAVE_API_KEY}" }
    },
    "postgres-db": {
      "command": "npx",
      "args": ["-y", "@modelcontextprotocol/server-postgres"],
      "env": {
        "PGHOST": "${PGHOST}",
        "PGPORT": "${PGPORT}",
        "PGDATABASE": "${PGDATABASE}",
        "PGUSER": "${PGUSER}",
        "PGPASSWORD": "${PGPASSWORD}"
      }
    },
    "memory": {
      "command": "npx",
      "args": ["-y", "@modelcontextprotocol/server-memory"]
    },
    "sequential-thinking": {
      "command": "npx",
      "args": ["-y", "@modelcontextprotocol/server-sequential-thinking"]
    },
    "large-file": {
      "command": "npx",
      "args": ["-y", "@willianpinho/large-file-mcp"]
    }
  }
}
\end{lstlisting}

\subsection{Deployment Instructions}
To apply this universal configuration, copy the JSON block above and place it in the respective configuration file for your AI client.

\begin{itemize}
    \item \textbf{Claude Desktop}: 
    \begin{itemize}
        \item \textit{macOS/Linux}: \lstinline[style=inline]|~/.config/Claude/claude_desktop_config.json|
        \item \textit{Windows}: \lstinline[style=inline]|%APPDATA%\Claude\claude_desktop_config.json|
    \end{itemize}
    
    \item \textbf{Gemini CLI}: 
    \begin{itemize}
        \item Place the JSON inside \lstinline[style=inline]|~/.gemini/settings.json| or a local \lstinline[style=inline]|.gemini/settings.json| at your project root.
        \item Verify successful bindings by running \lstinline[style=inline]|/mcp list| inside the prompt.
    \end{itemize}
    
    \item \textbf{Kiro IDE}:
    \begin{itemize}
        \item Map the configuration to \lstinline[style=inline]|~/.config/Kiro/mcp_config.json|.
    \end{itemize}
    
    \item \textbf{Cursor IDE}:
    \begin{itemize}
        \item While Cursor manages configurations via a UI, you can manually input these servers by navigating to \textbf{Settings -> Features -> MCP} and adding each entry using the ``command'' type matching the JSON definitions.
    \end{itemize}
\end{itemize}
