\documentclass{article}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{geometry}
\geometry{margin=1in}
\usepackage{enumitem}
\usepackage{hyperref}
\hypersetup{
  colorlinks=true,
  linkcolor=black,
  urlcolor=blue,
  citecolor=black
}
\usepackage{xcolor}
\usepackage{listings}
\definecolor{codebg}{gray}{0.96}
\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  columns=fullflexible,
  keepspaces=true,
  showstringspaces=false,
  frame=single,
  rulecolor=\color{gray!50},
  backgroundcolor=\color{codebg},
  framesep=6pt,
  framerule=0.4pt,
  aboveskip=0.6\baselineskip,
  belowskip=0.6\baselineskip
}
\lstdefinestyle{inline}{%
  basicstyle=\ttfamily\small,
  showstringspaces=false
}

% Bordered table with paragraph-like spacing above and below.
\newenvironment{paratable}{%
  \par\addvspace{\baselineskip}%
  \noindent
  \begingroup
  \renewcommand{\arraystretch}{1.4}%
}{%
  \endgroup
  \par\addvspace{\baselineskip}%
}






\title{The MCP Ecosystem\\[0.25em]\large Distribution \& Discovery}
\author{Jaehoon Song}
\date{\today}







 


\begin{document}
\maketitle

\textbf{Model Context Protocol (MCP)} is a standardized communication \textbf{protocol}
that securely connects AI large language \textbf{models} (LLMs) to external
\textbf{contexts} like codebases, databases, and files, frequently using either
standardized input and output (\textbf{STDIO}) for local integration or
Server-Sent Events (\textbf{HTTP/SSE}) for network-based remote communication.

This note explains how MCP \textbf{servers} are published, how you install them, and
how an AI \textbf{agent} learns what they can do, without you describing every
\textbf{tool} by hand.

\section{Package Distribution}

MCP servers are ordinary programs, usually written in \textbf{TypeScript (Node.js)} or
\textbf{Python}. They are distributed the same way as everyday \textbf{packages}:

\begin{itemize}
  \item \textbf{NPM}: run TypeScript servers on demand with \lstinline[style=inline]|npx|.
\begin{lstlisting}
npx -y @modelcontextprotocol/server-github
\end{lstlisting}
  \item \textbf{PyPI}: run Python servers with \lstinline[style=inline]|uvx| or \lstinline[style=inline]|pip|.
\begin{lstlisting}
uvx mcp-server-sqlite
\end{lstlisting}
\end{itemize}

Because servers are published packages, an AI \textbf{host} (such as Claude Desktop) can
download and launch the latest version on the fly, with no manual install step required.

\section{Official, Community, and Custom Servers}

The \textbf{protocol} is open to everyone, but servers fall into a few familiar tiers:
\begin{paratable}
\begin{tabular}{|l|p{0.68\textwidth}|}
\hline
\textbf{Official}\vspace{0pt} &
Maintained by the MCP team (Anthropic) and major partners.
Hosted at \url{https://github.com/modelcontextprotocol/servers}.
Reference implementations for services such as GitHub, Google Drive, and Slack.
\\
\hline
\textbf{Community}\vspace{0pt} &
Built by independent developers.
Discoverable on GitHub or in galleries such as
\url{https://mcp-get.com} and \url{https://smithery.ai}.
\\
\hline
\textbf{Private / Custom}\vspace{0pt} &
\textbf{Tools} you build for your own workflow: local scripts, internal APIs, or
project-specific utilities.
\\
\hline
\end{tabular}
\end{paratable}

\section{Automatic Tool Discovery}

How does the AI know what a server can do?

\begin{enumerate}
  \item \textbf{Handshake.}
    When the \textbf{host} starts an MCP server, it sends a \lstinline[style=inline]|tools/list| request
    (part of the \textbf{JSON-RPC} conversation between \textbf{client} and
    \textbf{server}).
  \item \textbf{Self-description.}
    The server replies with a structured \textbf{schema} for each tool: its name,
    \textbf{parameters}, and a human-readable \textbf{description} of what it does.
  \item \textbf{Reasoning.}
    The \textbf{model} reads that schema at runtime.
    If you ask \emph{``Is it raining in Tokyo?''} and a tool
    \lstinline[style=inline]|get_weather(city: string)| is listed, the model can match your question
    to the tool and call it.
\end{enumerate}

\textit{Example tool declaration (conceptual):}
\begin{quote}
  \lstinline[style=inline]|get_weather| needs \lstinline[style=inline]|city| (string).
  Returns local temperature and conditions for the given city.
\end{quote}

No manual prompt engineering is required for each tool; the server advertises its
\textbf{capabilities}, and the host exposes them to the model.

\section{Connecting Servers to Your AI Client}

Adding a new capability to your AI client follows the same pattern every time:

\begin{enumerate}
  \item \textbf{Find}: look up the package for the service you want
  (e.g.\ Brave Search, Postgres).
  \item \textbf{Auth}: obtain an API key or connection string for that service.
  \item \textbf{Config}: add the launch command to your host configuration
  (e.g.\ \lstinline[style=inline]|claude_desktop_config.json|).
  \item \textbf{Restart}: restart the AI client so it loads the new server.
\end{enumerate}


After that, the model has new \emph{hands}: structured \textbf{tools} it can invoke to
act on that part of the world on your behalf.

\end{document}
